Topological Equivariant Artist Model

### 2021-12-27 slides/main.tex
most of math section:

\usepackage{graphicx}

@@

\begin{frame}{Equivariance}
    \begin{description}
        \item[Retinal Variables \& Marks] visual encodings match properties of the data \cite{bertinSemiologyGraphicsDiagrams2011a}
        \pause
        \item[Expressiveness] structure preserving mappings\cite{mackinlayAutomatingDesignGraphical1986}
        \pause
        \item[Graphical Integrity] graphs show \textbf{only} the data\cite{tufteVisualDisplayQuantitative2001}
        \pause 
        \item[Naturalness] easier to understand when properties match\cite{norman_things_smart}
    \end{description}
\end{frame}

\begin{frame}{Equivariance}

\begin{frame}{Domain specific libraries assume data structure\cite{HeerSoftware2006}}
    \begin{table}
        %\renewcommand{\arraystretch}{2}
        %
        \begin{tabular}{>{\onslide<1->}l>{\onslide<2->}l>{\onslide<3->}l}
            \includegraphics[width=.24\textwidth]{figures/intro/table.png} & \includegraphics[width=.3\textwidth]{figures/intro/landsat.png} & \includegraphics[width=.33\textwidth]{figures/math/graph.png} \\
            ggplot\cite{wickhamGgplot2ElegantGraphics2016a}  & ImageJ\cite{schneiderNIHImageImageJ2012}& Gephi\cite{bastianGephiOpenSource2009}\\
            Vega\cite{satyanarayanDeclarativeInteractionDesign2014} & ImagePlot\cite{studiesCulturevisImageplot2021} & Graphviz\cite{ellsonGraphvizOpenSource2002}\\
            Altair\cite{vanderplasAltairInteractiveStatistical2018}& Napari\cite{nicholas_sofroniew_2021_4533308} & Networkx\cite{HagbergExploringNetwork2008}\\
             Tableau \cite{StoltePolaris2002}& &\\
            \cite{hanrahanVizQL2006,MackinlayShowme2007}&&\\

@@

\begin{frame}{General purpose libraries can't\cite{toryRethinkingVisualizationHighlevel2004}}
    \begin{figure}
        \includegraphics[height=.5\textheight]{figures/intro/dataset_diagram.png}
        \caption{Data Representation, MayaVi 4.7.2 docs\cite{DataRepresentationMayavi}}
    \end{figure}
    \begin{enumerate}
        \item Matplotlib\cite{hunterMatplotlib2DGraphics2007} $\rightarrow$Seaborn\cite{waskom2020seaborn}, xarray \cite{hoyer2017xarray} 
        \item D3 \cite{bostockDataDrivenDocuments2011}
        \item VTK \cite{hanwellVisualizationToolkitVTK2015,geveciVTK2012}(MayaVi\cite{RamachandranMayaVI2011})$\rightarrow$ Titan\cite{brianwylieUnifiedToolkitInformation2009}, ParaView\cite{ahrens2005paraview}
    \end{enumerate}
\end{frame}

@@

\begin{frame}{Fiber Bundles}
    \begin{columns}
        \column{0.49\columnwidth}
            \begin{equation*}
                %\label{eq:related-work:continuity:fiber-bundle}
                \begin{tikzcd}[ampersand replacement=\&]
                  \dfiberc \arrow[r, hook] \& \dtotalc \arrow[d, "\pi"'] \\
                   \& \dbasec \arrow[u, "\dsectionc \in \cgamma{\dbasec}{\dtotalc}"', bend right, pos=.5]
                \end{tikzcd}
            \end{equation*}
     \column{0.49\columnwidth}
        \begin{description}
            \item[\textcolor{base}{Base Space}] $( \dbasec, \mathcal{T})$
            \item[\textcolor{fiber}{Fiber Space}] $\dfiberc_{\dbasepointc} = \pi^{-1}(\dbasepointc)$
            \item[\textcolor{total}{Total Space}]  $\dtotalc = \dbasec \times \dfiberc$
        \end{description}
    \end{columns}
    \pause
    \begin{block}{Sections of the bundle}
        \begin{equation*}
            \label{eq:related-work:continuity:sections}
            \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} \coloneqq \big\{\dsectionc: \opensetc\rightarrow \dtotalc\restriction_{\openset} \; \bigm{\vert} \pi(\dsectionc(\dbasepointc)) = \dbasepointc\;\forall \dbasepointc \in \opensetc \subset \dbasec \big\}
        \end{equation*}
    \end{block}
\end{frame}
\subsection{Contributions}
\begin{frame}{Contributions}
    \begin{description}
        \item[Topological] continuity  
        \item[Equivariant] monoid action 
        \item[Artist] Matplotlib $\textit{artist}: \textit{data} \rightarrow \textit{graphic}$ 
        \item [Model] 
    \end{description}
\end{frame}

@@

\begin{frame}{Mathematical Models of Vizualization}
    \begin{description}
        \item[algebraic process] data and viz transforms are symmetric \cite{kindlmannAlgebraicProcessVisualization2014}
        \begin{equation*}
            \begin{tikzcd}[ampersand replacement=\&]
                \dtotal \arrow[d, "m"'] \arrow[r, "\nu"] \& \vtotal \arrow[r, "Q"]  \& \gtotal \arrow[d, "\varphi(m)"] \\
                \dtotal \arrow[r, "\nu"']                     \& \vtotal \arrow[r, "Q"'] \& \gtotal                  
            \end{tikzcd}
        \end{equation*}

        \item[language] APT and GoG: syntax, semantics, and grammar\\ 
        \cite{mackinlayAutomatingDesignGraphical1986,mackinlayAUTOMATICDESIGNGRAPHICAL1987,wilkinsonGrammarGraphics2005}
        \item[functional dependencies] relationship between components \cite{sugibuchiFramwork2009}
        \item[category theory]  \textit{understanding} = \textit{read} $\circ$ \textit{render} \cite{vickersUnderstandingVisualizationFormal2013}    
    \end{description}
\end{frame}

@@

    def view(self, axes):
        # walk the edge_vertex table to return the edge function
        table = defaultdict(list)
        for (i, (start, end)) in sorted(zip(self.ids, self.vertices), 
                                            key=lambda v:v[1][0]):
            table['index'].append(i)
            # same as view for line, returns nested list
            for (name, value) in zip(self.FB.F.keys(), self.tau(i)[1]):
                table[name].append(value)
        return table
\end{minted}
\end{frame}
%\end{comment}
\end{document}

>>>>

\usepackage{graphicx}
\usepackage{adjustbox}%scale tikz-cd

@@

\begin{frame}{Equivariance}
    \begin{description}
        \item[Retinal Variables \& Marks] visual encodings should match properties of the data \cite{bertinSemiologyGraphicsDiagrams2011a}
        \pause
        \item[Graphical Integrity] graphs show \textbf{only} the data\cite{tufteVisualDisplayQuantitative2001}
        \pause 
        \item[Naturalness] easier to understand when properties match\cite{norman_things_smart}
        \pause
        \item[Expressiveness] which structure preserving mappings can a tool implement\cite{mackinlayAutomatingDesignGraphical1986}] 
    \end{description}
\end{frame}

\begin{frame}{Frameworks for Expressing Visual Equivariance}
    \begin{description}
        \item[linguistically] visualization has syntax, semantics, grammar expresses how to design structure preserving visualizations
        \cite{mackinlayAutomatingDesignGraphical1986,mackinlayAUTOMATICDESIGNGRAPHICAL1987,wilkinsonGrammarGraphics2005}

        \item[algebraically] transformations on data and graphics are equivalent symmetric \cite{kindlmannAlgebraicProcessVisualization2014}
        \begin{columns}
            \column{.5\textwidth}
            \begin{description}
                \item[D] data 
                \item[R] data representation 
                \item[V] visualization
            \end{description}
            \column{.5\textwidth}
            \begin{equation*}
                \begin{tikzcd}[ampersand replacement=\&]
                    D \arrow[d, "\alpha"'] \arrow[r, "r_1"] \& R \arrow[r, "\nu"]  \& V \arrow[d, "\omega"] \\
                    D \arrow[r, "r_2"']                     \& R \arrow[r, "\nu"'] \& V                    
                \end{tikzcd}
                \end{equation*}
        \end{columns} 
    \item[categorically] \textit{understanding} = \textit{read} $\circ$ \textit{render} \cite{vickersUnderstandingVisualizationFormal2013}    
    \end{description}
\end{frame}

\begin{frame}{Domain specific libraries know their structure\cite{HeerSoftware2006}}
    \begin{table}
        %\renewcommand{\arraystretch}{2}
        %
        \begin{tabular}{>{\onslide<1->}l>{\onslide<2->}l>{\onslide<3->}l}
            \includegraphics[width=.24\textwidth]{figures/intro/table.png} & \includegraphics[width=.3\textwidth]{figures/intro/landsat.png} & \includegraphics[width=.33\textwidth]{figures/math/graph.png} \\
            ggplot\cite{wickhamGgplot2ElegantGraphics2016a}  & ImageJ\cite{schneiderNIHImageImageJ2012}& Gephi\cite{bastianGephiOpenSource2009}\\
            Vega\cite{satyanarayanDeclarativeInteractionDesign2014} & ImagePlot\cite{studiesCulturevisImageplot2021} & Graphviz\cite{ellsonGraphvizOpenSource2002}\\
            Altair\cite{vanderplasAltairInteractiveStatistical2018}& Napari\cite{nicholas_sofroniew_2021_4533308} & Networkx\cite{HagbergExploringNetwork2008}\\
             Tableau \cite{StoltePolaris2002}& &\\
            \cite{hanrahanVizQL2006,MackinlayShowme2007}&&\\

@@

\begin{frame}{General purpose libraries generally can't\cite{toryRethinkingVisualizationHighlevel2004}}
    \begin{figure}
        \includegraphics[height=.3\textheight]{../paper/figures/k_different_types.png}
    \end{figure}
    \begin{enumerate}
        \item Matplotlib\cite{hunterMatplotlib2DGraphics2007} $\rightarrow$Seaborn\cite{waskom2020seaborn}, xarray \cite{hoyer2017xarray} 
        \item D3 \cite{bostockDataDrivenDocuments2011}
        \item VTK \cite{hanwellVisualizationToolkitVTK2015,geveciVTK2012}(MayaVi\cite{RamachandranMayaVI2011})$\rightarrow$ Titan\cite{brianwylieUnifiedToolkitInformation2009}, ParaView\cite{ahrens2005paraview}
    \end{enumerate}
\end{frame}

@@

\begin{frame}{Fiber Bundles}
    \begin{columns}
        \column{0.39\textwidth}
            \begin{equation*}
                %\label{eq:related-work:continuity:fiber-bundle}
                \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
                  \dfiberc \arrow[r, hook] \& \dtotalc \arrow[d, "\pi"'] \\
                   \& \dbasec \arrow[u, "\dsectionc \in \cgamma{\dbasec}{\dtotalc}"', bend right, pos=.5]
                \end{tikzcd}
            \end{equation*}
     \column{0.59\textwidth}
        \begin{description}
            \item[\textcolor{base}{Base Space}] $( \dbasec, \mathcal{T})$ with
                \begin{itemize}
                   \item open sets $\opensetc \subset \dbasec$
                   \item points $\dbasepointc \in \opensetc \in \dbasec$
                 \end{itemize}  
            \item[\textcolor{fiber}{Fiber Space}] $\dfiberc_{\dbasepointc} = \pi^{-1}(\dbasepointc)$
            \item[\textcolor{total}{Total Space}]  $\dtotalc\restriction_{\opensetc} = \dbasec\restriction_{\opensetc} \times \dfiberc\restriction_{\opensetc}$ \\ for all $\opensetc \subset K$ 
        \end{description}
    \end{columns}
    \pause
    \begin{alertblock}{Sections}
    $\cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} \coloneqq \big\{\dsectionc: \opensetc\rightarrow \dtotalc\restriction_{\openset} \; \bigm{\vert} \pi(\dsectionc(\dbasepointc)) = \dbasepointc\;for\, all\; \dbasepointc \in \opensetc \big\}$
    \end{alertblock}
\end{frame}

\begin{frame}{Category $\textcolor{source}{\mathcal{C}}$}
\begin{columns}
    \column{0.39\textwidth}
\begin{equation*}
    \begin{tikzcd}[ampersand replacement=\&]
        \textcolor{source}{C_1} \arrow[r, "f", color=source] \arrow[rd, "g\circ f"', color=source] \arrow["id_{C_1}"', loop, distance=2em, in=215, out=145, color=source] \& \textcolor{source}{C_2} \arrow[d, "g", color=source] \arrow["id_{C_2}"', loop, distance=2em, in=125, out=55, color=source] \\
    \& \textcolor{source}{C_3} \arrow["id_{C_3}"', loop, distance=2em, in=305, out=235, color=source]              
    \end{tikzcd}
    \end{equation*}
    \column{0.59\textwidth}
    \begin{alertblock}{associativity} 
        if $f: C_1 \rightarrow C_2$, $g: C_2 \rightarrow C_3$ and $h: C_3 \rightarrow C_4$ then $h\circ (g \circ f) = (h \circ g) \circ f$
    \end{alertblock}
    \begin{alertblock}{identity} 
        for every $f: C_1 \rightarrow C_2$ there exists identity morphisms $f \circ id_{C_1} = f = id_{C_2} \circ f$
    \end{alertblock}
    \end{columns}
\end{frame}

\begin{frame}{Opposite Category $\textcolor{source}{\mathcal{C}^{op}}$}
    \begin{columns}
        \column{0.49\textwidth}
    \begin{equation*}
        \begin{tikzcd}[ampersand replacement=\&]
            \textcolor{fade}{C_1} \arrow[r, "f", color=fade] \arrow[rd, "g\circ f"', color=fade] \arrow["id_{C_1}"', loop, distance=2em, in=215, out=145, color=fade] \& \textcolor{fade}{C_2} \arrow[d, "g", color=fade] \arrow["id_{C_2}"', loop, distance=2em, in=125, out=55, color=fade] \\
        \& \textcolor{fade}{C_3} \arrow["id_{C_3}"', loop, distance=2em, in=305, out=235, color=fade]              
        \end{tikzcd}
        \end{equation*}
        \column{0.49\textwidth}
        \begin{equation*}
        \begin{tikzcd}[ampersand replacement = \&]
            \textcolor{source}{C_1} \arrow["id_{C_1}"', loop, distance=2em, in=215, out=145, color=source] \& \textcolor{source}{C_2} \arrow[l, "f"',color=source] \arrow["id_{C_2}"', loop, distance=2em, in=125, out=55, color=source]\\
            \& \textcolor{source}{C_3} \arrow[u, "g"', color=source] \arrow[lu, "f \circ g", color=source] \arrow["id_{C_3}"', loop, distance=2em, in=305, out=235, color=source]
            \end{tikzcd}
        \end{equation*}
        \end{columns}
\end{frame}

\begin{frame}{Functor $\textcolor{sheaf}{\bm{F}}: \textcolor{source}{\mathcal{C}} \rightarrow \textcolor{target}{\mathcal{D}}$}
\begin{columns}
    \column{.39\textwidth}
    \begin{equation*}
        \begin{tikzcd}[ampersand replacement = \&, ]
            \textcolor{source}{c} \arrow[d, "f"', shift right=0, color=source] \arrow[r, "F", color=functor] \& \textcolor{target}{F(c)} \arrow[d, "F(f)", shift right=0, color=target] \\
            \textcolor{source}{c^{\prime}} \arrow[r, "F"', color=functor]                       \& \textcolor{target}{F(c^{\prime})}                      
        \end{tikzcd}
    \end{equation*}
    \column{.59\textwidth}
    \begin{alertblock}{composition}
    $\textcolor{functor}{F}(\textcolor{source}{g}) \circ  \textcolor{functor}{F}(\textcolor{source}{f}) = \textcolor{functor}{F} (\textcolor{source}{g}\circ \textcolor{source}{f})$
    \end{alertblock}
    \begin{alertblock}{identity}
    $\textcolor{functor}{F}(\textcolor{source}{id_c}) = \textcolor{target}{id}_{\textcolor{functor}{F}(\textcolor{source}{c})}$
    \end{alertblock}
\end{columns}
\end{frame}

\begin{frame}{Natural Transformation $\textcolor{nattran}{\alpha}: \textcolor{functor}{F} \textcolor{nattran}{\rightarrow} \textcolor{functor}{G}$}

\adjustbox{scale=3, center}{
    \begin{tikzcd}[ampersand replacement=\&]
        \& {} \arrow[dd, "\alpha", Rightarrow, shorten=1.5em, color=nattran] \&             \\
      \textcolor{source}{\mathcal{C}} \arrow[rr, "F", bend left, color=functor] \arrow[rr, "G"', bend right, color=functor] \& \& \textcolor{target}{\mathcal{D}} \\
        \& {} \&            
      \end{tikzcd}}
\end{frame}

\begin{frame}{Natural Transformation $\textcolor{nattran}{\alpha}: \textcolor{functor}{F} \textcolor{nattran}{\rightarrow} \textcolor{functor}{G}$}
\begin{equation*}
    \begin{tikzcd}[column sep=huge, ampersand replacement = \&]
        \textcolor{target}{F(c)} \arrow[dd, "F(f)"', color=target] \arrow[rr, "\alpha_{c}", bend left, color=nattran] \& \textcolor{source}{c} \arrow[dd, "f" description, color=source] \arrow[r, "G"', color=functor] \arrow[l, "F", color=functor] \& \textcolor{target}{G(c)} \arrow[dd, "G(f)", color=target] \\
        {} \arrow[rr, dotted, color=nattran] \& \& {}                      \\
        \textcolor{target}{F(c^{\prime})} \arrow[rr, "\alpha_{c^{\prime}}"', bend right, color=nattran] \& \textcolor{source}{c^{\prime}} \arrow[l, "F"', color=functor] \arrow[r, "G", color=functor] \& \textcolor{target}{G(c^{\prime})}          
    \end{tikzcd}
\end{equation*}    
\end{frame}

\begin{frame}{Presheaf: $\sheafc: \opensetc \textcolor{sheaf}{\rightarrow} \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}$}
\begin{equation*}
    \begin{tikzcd}[ampersand replacement = \&, column sep=small]
        {\cgamma{\opensetc_1}{\dtotalc\restriction_{{\opensetc_1}}}} \& {} \& {\cgamma{\opensetc_2}{\dtotalc\restriction_{\opensetc_2}}} \arrow[ll, "\iota^*"', hook', color=set] \\\& \&\\
        \opensetc_1 \arrow[uu, "\sheafc_{\dbasec}", color=sheaf] \arrow[rr, "\iota"', hook, color=base] \& {} \arrow[uu, "\sheafc_{\dbasec}" description, dotted, color=sheaf] \& \opensetc_2 \arrow[uu, "\sheafc_{\dbasec}"', color=sheaf]                    
        \end{tikzcd}
\end{equation*}
\end{frame}

\begin{frame}{Sheaf}

\end{frame}
\begin{frame}{Stalk \& Germ}

\adjustbox{scale=1}{
    \begin{description}
    \item[Stalk] $\sheafc(\dbasec)\restriction_{\dbasepointc}\coloneqq \lim\limits_{\opensetc\ni \dbasepointc} \Gamma(\opensetc, \dtotalc\restriction_{\opensetc})$
    \item[Germ] $\dsectionc(\dbasepointc) \in \sheafc(\dbasec)$
\end{description}
}
\end{frame}

\begin{comment}
\subsection{Contributions}
\begin{frame}{Contributions}
    \begin{description}
        \item[Topological] continuity  
        \item[Equivariant] monoid action 
        \item[Artist] Matplotlib $\textit{artist}: \textit{data} \rightarrow \textit{graphic}$ 
        \item [Model] 
    \end{description}
\end{frame}

@@

\begin{frame}{Bertin Retinal Variables}
    \begin{figure}
        \includegraphics[width=.7\textwidth]{figures/intro/retinal_variables.png}
        \caption{This tabular form of Bertin's retinal variables is from Understanding Graphics \cite{malamedInformationDisplayTips2010} who reproduced it from Krygier and Wood's \textit{Making Maps: A Visual Guide to Map Design for GIS}\cite{krygierMakingMapsVisual2005}}
    \end{figure}
\end{frame}

@@

    def view(self, axes):
        # walk the edge_vertex table to return the edge function
        table = defaultdict(list)
        for (i, (start, end)) in sorted(zip(self.ids, self.vertices), 
                                            key=lambda v:v[1][0]):
            table['index'].append(i)
            # same as view for line, returns nested list
            for (name, value) in zip(self.FB.F.keys(), self.tau(i)[1]):
                table[name].append(value)
        return table
\end{minted}
\end{frame}
%\end{comment}

\end{comment}
\end{document}
